---
title: "update_azmpdata"
author: Emily Chisholm
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{update_azmpdata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(azmpdata)
```


This package should be updated on an annual basis or whenever there are major changes made to the data set. The update process should only be undertaken by developers or data managers. The following document describes the procedure for updating the data package


### Step 1

__Add new raw data files__

For physical data: Raw data files (in either .dat or .csv) can be added to `inst/extdata`. It is best if this folder is kept organized by variable, for example the `inst/extdata/airTemperature` folder contains all the raw airTemperature data. It is important that the raw data format be identical to the previously used formats for that specific variable. The format does vary between variables but should be kept the same between years within a variable. If you are unsure of the correct format, please open an existing data file and review the format before uploading new raw data!

For biological data: Raw data files are uploaded to the FTP server at 'ftp://ftp.dfo-mpo.gc.ca/AZMP_Maritimes/AZMP_Reporting/outputs'. The formatting of these data files is important and should be maintained between years.

For more details on the formatting of data, please see the `style_guide` vignette. 

### Step 2
__Processing Scripts__
Data is loaded into the package using individual scripts for each dataframe. These scripts are contained in `inst/extdata` and organized by folder depending on the data type. These scripts output package data objects. 

### Step 3
__Source all scripts__
To recreate all dataframes use the R script 'source_all.R' this will regenerate all dataframes in the package and update the datadate.txt file which identifies the data on which data was last updated. This datadate.txt file is important for versioning the package and ensuring users have the latest data. If running dataframe scripts individually, please ensure datadate.txt is updated using the `data_update()` function. 

