---
title: "Meeting Notes - June 10"
output: html_document
---
Discussion of Maritimes environmental data access, June 10, 2020

B. Casault, E. Chisholm, C. Layton, C. Johnson, M. McMahon

## Emily presented spreadsheet
* This spreadsheet contains all required multivaR metrics as well as all metrics provided by Benoit and Chantelle in reporting
* Details by category and variable
* an 'x' in column azmpdata means that it is a multivar metric to be included in azmpdata, a '*' means another reporting metric (not required for multivaR)

```{r}
library(xlsx)

f <- read.xlsx('supp_files/azmpmetrics.xlsx', sheetIndex = 1)

head(f)
```


* first part of the meeting was used to discuss metrics which are required multivaR but not included in azmpdata
* Chantelle has made some notes in spreadhseet about physical variables she is able to include in her outputs

* St. Lawrence river flux
  * should be included in azmpdata
  * will require annual confirmation of URL
  

* Ice volume
  * Included in Peter's products
  * Peter produces SS metrics and GSL metrics, multivaR uses SS + GSL metrics
  * __EC will investigate difference between SS+GSL and GSL, to determine if GSL metrics could be used as a valid proxy__

* Zooplankton metrics are all included

* SST warming metric will be included in azmpdata
  * BC has script to produce this
  * other option would be to put raw SST data in azmpdata and use BC script in multivaR to produce warming metric
  * CL is producing SST by NAFO region (monthly and annual values)
  * files come from Peter
  * CJ leans toward keeping warming metric directly in azmpdata
  
* Final Ice Day
  * __EC will invetsigate difference between SS+GSL and GSL__
  * Determine if it is worth combining before inclusion in azmpdata or if can be input straight from Peter
  
* Peak in C. fin biomass & peak in C fin (I, II, III) proportion
  * staged abundance will be included in azmpdata
  * proportion of I, II, III or peak biomass to be calculated in multivaR
  * CJ mention of hyperboreus abundance being included?
  
  
## Presentation from Chantelle 


* IDENTIFIED WHERE WATER IS SAMPLED VS AVERAGING DEPTHS FOR PHYSICAL DATA
  * Not all physical avergaing depths align with biological sampling depths
  * specific issues at P5 and AZMP transects
  
  
```{r}


f <- read.xlsx('supp_files/azmpDataFileColumns.xlsx', sheetIndex = 1)

head(f)

```


```{r}


f <- read.xlsx('supp_files/annualReportingProducts.xlsx', sheetIndex = 1)

head(f)
```
* above data table shows potential groupings of similarly scaled data for insertion into azmpdata
* Discussion about depth resolved data in azmpdata
  * Cons:
    * Large file size
    * Repetitive of BioChem and other DBs
  * Pros
    * solves Benoit's problems when data is requested
    * Eases difficulty accessing BioChem
  * should be disscussed with ODIS
  * spatial data will be in new DB being created by Gordana
  * azmpdata should focus on calculated metrics and variables?
  * could also be seperated into a distinct data package
  * current plan: to include depth resolved data only at bottle depths
  

## Presentation from Benoit


* Presented slides
__ADD IMAGE__



* Seperated data into groups of files
* core and fixed stations are ditinct
* CJ: zooplankton should be seperated carefully by group to avoid accidental summation of duplicated data in one file
* would have 12 core stations files using this plan
* naming scheme suggestions?
* variables need consistent naming scheme
* anomoly inclusion
  * BC: No, due to changes in calculations and reference periods
  * keep data in azmpdata and allow users to calculate anomolies
  * __CJ: EC should make anomoly calculations in a seperate package__
* BC: Goal to create another data package for remote sensing data
* Include fixed station metadata files created by BC
* Accessory files for polygons etc

## Meeting will pick back up with discussion on materials from Benoit, June 24th @ 2pm




